The goal of homework 3 is to test your understanding of Energy-Based Models, and to show you one application in structured prediction.

In the theoretical part, we'll mostly test your intuition. You'll need to write brief answers to questions about how EBMs work. In part 2, we will implement a simple optical character recognition system. 

In part 1, you should submit all your answers in a pdf file. As before, we recommend using \LaTeX. 

For part 2, you will implement some neural networks by adding your code to the provided ipynb file.

As before, please use numerator layout.

The due date of homework 3 is 5:00pm \texttt{03/25}.
Submit the following files in a zip file \texttt{your\_net\_id.zip} through NYU classes:
\begin{itemize}
\item \texttt{hw3\_theory.pdf}
\item \texttt{hw3\_practice.ipynb}
\end{itemize}

\textbf{Note: we will subtract points for Campuswire posts containing solutions to problems.} Campuswire shouldn't be a platform where you can get your solution checked, the goal is to help you with any misunderstandings associated with the homework.

The following behaviors will result in penalty of your final score:
\begin{enumerate}
\item 10\% penalty for submitting your file without using the correct naming format (including naming the zip file, PDF file or python file wrong, adding extra files in the zip folder, like the testing scripts in your zip file). 
\item 20\% penalty for late submission within the first 24 hours after the deadline. We will not accept any late submission after the first 24 hours.
\item 20\% penalty for code submission that cannot be executed following the steps we mentioned.
\end{enumerate}